---
title: "Graph Customization"
author: "Aiste Guobyte"
format: 
  html:
    toc: true
    code-fold: false
    code-summary: "Show code"
date: "2026-05-13"
brand: _brand.yml
---

## Intro

This document summarizes helpful tips for simple graph customization.

```{r, warning = F, message = F}
library(tidyverse)
```


## Create a Theme

First, set your colors for easy access. Color schemes can be created using online palette generators as well as packages like [MonochromeR](https://github.com/cararthompson/monochromeR) by Cara Thompson. Here is a link to the [ShinyApp](https://cararthompson.shinyapps.io/monochromeR/).

Ensure your color palettes are accessible for people with color blindness.

```{r}

cred   <- "#AB1721"
cblue  <- "#2A3B6B"
cdark  <- "#1A1A2E"
cwhite <- "#FFFFFF"

```

Next, create a chunk of code that includes edits to a specific theme (e.g., theme_bw). This can be pasted on the start of a document to streamline all charts in the document. Play around with your set color scheme for uniformly themed but different colored graphs throughout a document. Edits for each part of the graph can be found [here](https://ggplot2.tidyverse.org/reference/theme.html).

```{r}

list(theme_bw(base_size = 15, base_family = "Roboto Mono"), # use theme_bw to check default settings of theme bw.
  labs(x = NULL), # remove this if the x label should appear 
  theme(panel.grid.minor.x = element_blank(), # only y major remains 
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.title = element_text(color = cred, hjust = 0.5), # centered title
    text = element_text(color = cdark), # set color for all text
    axis.text.y = element_text(angle = 0, hjust = 1, color = cdark),
    axis.text.x = element_text(color = cdark),
    axis.line.x = element_line(linewidth = 1, color = cdark),
    axis.line.y = element_blank(), # left only the bottom line
    panel.border = element_blank(), # no border
    strip.background = element_blank(),
    strip.placement =  "outside",
    panel.background = element_blank())) -> CSimpleTheme # no border
    
```

## Fonts

If you download new fonts and have trouble with your computer recognizing them (e.g., paths on local user rather than Windows/Fonts), you can use the package extrafont to import fonts into R and then load fonts to use them. This has worked quite well for me. 

```{r, include = F}

# extrafont::font_import() run if necessary to add new fonts
extrafont::loadfonts(device = "win")

```


## Example

Here is a graph using a simulated data set.

```{r, warning = F}
read_rds("tbi_long.rds") -> long_df

ggplot(long_df, aes(x = time, y = value, group = id)) +
  geom_line(alpha = 0.3, linewidth = 0.4, position = position_jitter(width = 0.05)) +
  stat_summary(
    aes(group = 1),
    fun = mean,
    geom = "line",
    color = "red",
    linewidth = 1.2) +
  labs(
    y = "Outcome",
    title = "Individual & Mean Change from Pre to Post") +
  scale_x_discrete(
    labels = c(gcs_score_post = "GCS Score Post",
               gcs_score_pre = "GCS Score Pre"),
    expand = expansion(mult = c(.15, .15))) + # fills in space left and right
  scale_y_continuous(limits = c(0, 15)) + 
  CSimpleTheme
```

## Other

I use a _brand.yml doc to set logos, fonts, and colors on my HTML docs. Credit goes to [Charlotte Wickham](https://github.com/cwickham/arcadonia-brand) who presented this at R/Medicine 2026.
